---
title: "ST 558 - Summer 2020 - Project 2"
author: "Yun Ji"
date: "7/3/2020"
output:
  rmarkdown::github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r import, include = FALSE}
library(tidyverse)
library(caret)
library(e1071)
library(randomForest)
library(gbm)
library(knitr)
library(rmarkdown)
seed <- 1
```

## Introduction
The purpose of this vignette is to compare the prediction accuracy of two types of regression models on a provided data set. The data used is an online news popularity data set for web articles published by Mashable over a two-year period. Models fitted to this data set consist of a multiple linear regression model and a nonlinear Random Forest regression model. The data description and download links can be [found here](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity#).

## Build Automation
I use the following code to generate data and fitted models for each day of the week. The `apply` function calls another R Markdown file (`YunJi_Project2_DayOfWeekModel.Rmd`) which contains the code needed to import, explore, clean, filter, split, and model the news popularity data for each day.

```{r automation, results = "hide", cache = TRUE}
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
output_file <- paste0(days, "Analysis.md")
params <- lapply(days, FUN = function(x){list(day = x)})
reports <- tibble(output_file, params)

apply(reports, MARGIN = 1,
      FUN = function(x){
        render(input = "YunJi_Project2_DayOfWeekModel.Rmd", 
               output_file = x[[1]], 
               params = x[[2]]
              )
      }
)
```

## Day of Week Models

The model vignettes for each day of the week are linked below. Each day's vignette contains data exploration, along with model fitting on the training data set and model evaluation for predicting the testing data set.

* [Monday Models](MondayAnalysis.html)

* [Tuesday Models](TuesdayAnalysis.html)

* [Wednesday Models](WednesdayAnalysis.html)

* [Thursday Models](ThursdayAnalysis.html)

* [Friday Models](FridayAnalysis.html)

* [Saturday Models](SaturdayAnalysis.html)

* [Sunday Models](SundayAnalysis.html)


