---
title: "ST 558 - Summer 2020 - Project 2"
author: "Yun Ji"
date: "7/3/2020"
output:
  rmarkdown::github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r import, include = FALSE}
library(tidyverse)
library(caret)
library(e1071)
library(randomForest)
library(gbm)
library(knitr)
library(rmarkdown)
seed <- 1
```

## Introduction
The purpose of this vignette is to compare the prediction accuracy of two types of regression models on a provided data set. The data used is an online new popularity data set for web articles published by Mashable over a two-year period. The data description and download links can be [found here](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity#).

## Build Automation
I use the following code to generate data and fitted models for each day of the week.

```{r automation, results = "hide"}
days <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
output_file <- paste0(days, "Analysis.md")
params <- lapply(days, FUN = function(x){list(day = x)})
reports <- tibble(output_file, params)

apply(reports, MARGIN = 1,
      FUN = function(x){
        render(input = "YunJi_Project2_DayOfWeekModel.Rmd", 
               output_file = x[[1]], 
               params = x[[2]]
              )
      }
)
```

## Day of Week Models

The Markdown files for each day of the week are linked below.

[Monday Models](MondayAnalysis.md)
[Tuesday Models](TuesdayAnalysis.md)
[Wednesday Models](WednesdayAnalysis.md)
[Thursday Models](ThursdayAnalysis.md)
[Friday Models](FridayAnalysis.md)
[Saturday Models](SaturdayAnalysis.md)
[Sunday Models](SundayAnalysis.md)


